---
title: 'Bioconductor Developer Day, 2016: Shiny Gadgets and Interactive Base Graphics'
author: "Sean Davis"
date: "`r Sys.Date()`"
output: revealjs::revealjs_presentation
vignette: >
  %\VignetteIndexEntry{Overview of Shiny Gadgets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Shiny

- interactively communicating _ideas_ and _results_
    - plots
    - data
    - model fitting
- for full application development
    - interact with data sources
    - collect and react to input
    - ...

---
  
## Gadgets vs Apps

- Gadgets are interactive tools that enhance the developer analyst experience
- Apps represent the _output_ of an analsys, Gadgets are for using during an analysis
- The audience for an App is the *end user*, while the audience for a Gadget is an *R user*
- Gadget fundamentals are shiny, but design is focused
    - on limited screen space
    - on useful functionality *during* analysis
    - performing a function or returning results
- Apps run on Shiny Server, while Gadgets are intended to be invoked in code

---

## Gadgets

Shiny Gadgets could be created for most any task you might do during data loading, cleaning, manipulation, and visualization.

- A tool to preview regular expressions for find/replace (sub and gsub)
- Visual selection tools for subsetting or outlier exclusion
- An easy-to-use UI for downloading data from a complicated API and turning it into a data frame
- ...

---

## Gadget structure

```{r gadgetSkeleton,eval=FALSE}
library(shiny)
library(miniUI)

myGadgetFunc <- function(inputValue1, inputValue2) {
  ui <- miniPage(
    ## Define layout, inputs, outputs, etc.
  )
  
  server <- function(input, output, session) {
    # Define reactive expressions, outputs, etc.
    # And then 
    stopApp(returnValue)
  }
  runGadget(ui, server)
}
```
  
## Gadget UI block
  
```{r miniUI,eval=FALSE}
  ui <- miniPage(
    gadgetTitleBar("My Gadget"),
    miniContentPanel(
      # Define layout, inputs, outputs
    )
  )
```

---

## Gadget Server block

```{r miniUI2,eval=FALSE}
  server <- function(input, output, session) {
    # Define reactive expressions, outputs, etc.

    # When the Done button is clicked, return a value
    observeEvent(input$done, {
      returnValue <- ...
      stopApp(returnValue)
    })
  }
```

## regex test example

```{r regex1,eval=FALSE}
regexTest = function(pattern="night", 
                     x = "We can turn day into night with this Gadget",
                     replace = "day") {
  
  ui = miniPage(
    gadgetTitleBar("Basic gsub tester"),
    miniContentPanel(
      textInput("text","Text:", x),
      textInput('pattern','Pattern to replace:', pattern),
      textInput("replacement","Text to substitute", replace),
      textOutput("out")
    )
  )
  
  server = function(input, output, session) {
    output$out = renderText( gsub(pattern = input$pattern,
                                 replace = input$replacement, 
                                 x = input$text) )
    observeEvent(input$done, {
      returnValue <- input$pattern
      stopApp(returnValue)
    })
  }
  runGadget(ui, server)
}
```

---

## ggbrush example

```{r brushPoints,eval=FALSE}
ggbrush <- function(dframe, xvar, yvar, viewer=paneViewer()) {
  
  ui <- miniPage(
    gadgetTitleBar("Drag to select points"),
    miniContentPanel(
      # The brush="brush" argument means we can listen for
      # brush events on the plot using input$brush.
      plotOutput("plot", height = "100%", brush = "brush")
    )
  )
  
  server <- function(input, output, session) {
    
    # Render the plot
    output$plot <- renderPlot({
      # Plot the data with x/y vars indicated by the caller.
      ggplot(dframe, aes_string(xvar, yvar)) + geom_point()
    })
    
    # Handle the Done button being pressed.
    observeEvent(input$done, {
      # Return the brushed points. See ?shiny::brushedPoints.
      stopApp(brushedPoints(dframe, input$brush))
    })
  }

  runGadget(ui, server)
}
```
